---
title: "Linear Mixed-Effects Models"
author: "Hossameldin Mohammed"
math: true
date: "2021-09-19"
output: html_document
---

```{r}
suppressPackageStartupMessages(library(AER))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(broom))
suppressPackageStartupMessages(library(nlme))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(lmerTest))
```

# Introduction

Basic regression models are fitted with a sample of $n$ *independent elements*. Given a set of $p$ regressors $X_{i,j}$ and a continuous response $Y_i$, we fit a model

$$Y_i = \beta_0 + \beta_1 X_{i,1} + \beta_2 X_{i,2} + \ldots + \beta_p X_{i,p} + \varepsilon_{i} \; \; \; \; \text{for} \; i = 1, \ldots, n$$


The coefficients $\beta_0, \dots, \beta_p$ are fixed and constant for **all the observed values** $\left(x_{i,1}, \dots, x_{i,p}, y_i\right)$.

These coefficients are called **fixed effects**. It is of our interest to evaluate whether they are statistically significant or not on the response.

### 2.1. Grunfeld's Investment Dataset

Consider the following example: to study how gross investment depends on the firm's value and capital stock, [Grunfeld (1958)](https://ezproxy.library.ubc.ca/login?url=https://search.proquest.com/docview/301899386?pq-origsite=summon) collected data from eleven different companies over the years 1935-1954.

> The data frame `Grunfeld` contains 220 observations from a balanced panel of 11 firms from 1935 to 1954 (20 observations per `firm`). The dataset includes a continuous response `investment` subject to two explanatory variables, `market_value` and `capital`.

Firstly, we will load the data which has the following variables:

- `investment`: the gross investment in millions of dollars (additions to plant and equipment along with maintenance), a continuous response.
- `market_value`: the firm's market value in millions of dollars, a continuous explanatory variable.
- `capital`: stock of plant and equipment in millions of dollars, a continuous explanatory variable.
- `firm`: a nominal explanatory variable with eleven levels indicating the firm (`General Motors`, `US Steel`, `General Electric`, `Chrysler`, `Atlantic Refining`, `IBM`, `Union Oil`, `Westinghouse`, `Goodyear`, `Diamond Match`, and `American Steel`).
- `year`: the year of the observation (it will not be used in our analysis).

```{r}
data(Grunfeld)
Grunfeld <- Grunfeld %>% rename(investment = invest, market_value = value)
head(Grunfeld)
tail(Grunfeld)
```

